The period during which I wrote this thesis was eventful, to say the least.
From living through the COVID-19 pandemic and the Netherlands' first national curfew since World War II,
to actually contracting COVID-19,
to dislocating my shoulder (twice), to my first paper rejection, to my subsequent first paper acceptance.
However, for almost a year, regularly working on my thesis seemed to be the only reliably constant aspect of my life.
And luckily so. 
Because despite it being arguably the most challenging academic endeavor I've undertaken, I had a great time doing it.

The excellent supervision I received from Sandro, Raquel, and Janie played an essential role in making this experience as enjoyable as it was.
Their contribution to my work has been invaluable.
The seemingly endless number of virtual meetings we had every Tuesday at 12:00 rarely failed to exceed the predetermined one-hour mark, and I always came out of them filled with new ideas and an insatiable thirst for coffee.
I owe special thanks to Sandro for being so readily available to discuss practically everything I threw at him (even though I could've sometimes thought my questions through some more). 
And I am especially grateful for the extensive feedback they consistently gave on my work, which was, admittedly, at times a lot more than I expected (I believe Sandro holds the record at 145 comments on 1.5 chapters).
Producing a submission-ready paper with them was also a very valuable and rewarding experience.
Moreover, I am grateful for them continuing to encourage me to do so, despite my initial hesitancy.
% I learned from them that doing research is hard, and requires attention to detail and
% Happy to see them in real life

Finally, I'd also like to express my gratitude towards my parents, Harry and Cyrillia, my brothers, Karel and Max, Meve, Emile, Michael, Philip, Morena, Joris, Floris, Floris, Coen, Bram, Najib, Haye, Isadora, Margot, Marios, Spela, David, Thomas, and Linnea.


% \bigskip
% \bigskip

\begin{flushright}
    Lennert Jansen \\
    \noindent \today \\
    \noindent Amsterdam, the Netherlands
\end{flushright}