

\begin{longtable}{| c | c | m{10cm} |}
\caption{Timetable for my planning on experimenting, evaluating, and writing.}
\label{tab:plan}\\
    % \begin{tabular}{| c | c | m{10cm} |}
        \hline
        \textbf{Month} & \textbf{Week} & \textbf{Task} \\
        \hline
        June-July & 26 & \begin{itemize} 
                            \item 28/06 Thesis proposal deadline. 
                            \item Pre-summer RPA meeting roundup.
                            \item Migrate from AWS EC2 to Lisa GPU.
                            \item Construct properly representative most frequently used words per age-group (unique to age-group).
                            \item First experiments with GPT-2 + BNC Age discriminator.
                        \end{itemize} \\
        \hline
        July & 27 & \begin{itemize}
                        \item  Break.
                    \end{itemize}\\
        \hline
        July & 28 & \begin{itemize}
                        \item Break.
                    \end{itemize}\\
        \hline
        July & 29 & \begin{itemize}
                        \item Coding for BNC-discrim. + GPT-2
                        \item Coding for BNC-discrim. + DialoGPT
                        \item Latent space and feature inspections age-discriminators.
                        \item PPCM code setup.
                    \end{itemize}\\
        \hline
        July-August & 30 &  \begin{itemize}
                                \item Coding for BAC-discrim + GPT-2
                                \item Coding for BAC-discrim + DialoGPT
                                \item Automatic evaluation age-discrim. + LMs.
                                \item Write short summaries and interpretations of results.
                            \end{itemize}\\
        \hline
        August & 31 & \begin{itemize}
                        \item Finalize writing of age-discrim. + LM results up until now.
                        \item Break from Friday 06/08 to Monday 09/08.
                      \end{itemize}\\
        \hline
        August & 32 & \begin{itemize}
                                \item Coding for user-ready adaptive system (PPCM).
                                \item Writing for results and discussion until now.
                      \end{itemize}\\
        \hline
        August & 33 & \begin{itemize}
                            \item Coding for user-ready adaptive system (PPCM).
                            \item Writing for results and discussion until now.
                       \end{itemize}\\
        \hline
        August & 34 & \begin{itemize}
                        \item Test-runs for user-ready adaptive system (PPCM).
                        \item Writing for results and discussion until now.
                      \end{itemize}\\
        \hline
        August-September & 35 & \begin{itemize}
                                    \item Test-runs for user-ready adaptive system (PPCM).
                                    \item Writing for results and discussion until now.
                                \end{itemize}\\
        \hline
        September & 36 &\begin{itemize}
                                \item First RPA meeting after summer. Set up plans for next phase of experiments.
                        \end{itemize}\\
        \hline
        September & 37 & \begin{itemize}
                                \item Phase 2 user experiments.
                                \item Write tentative results.
                            \end{itemize}\\
        \hline
        September & 38 & \begin{itemize}
                                \item Phase 2 user experiments.
                                \item Write tentative results.
                        \end{itemize}\\
        \hline
        September-October & 39 &\begin{itemize}
                                    \item 01/10 Submit draft of final version for feedback.
                                \end{itemize}\\
        \hline
        October & 40 & \begin{itemize}
                            \item Time for additional experiments and case studies.
                        \end{itemize}\\
        \hline
        October & 41 & \begin{itemize}
                                \item Time for additional experiments and case studies.
                        \end{itemize}\\
        \hline
        October & 42 & \begin{itemize}
                            \item Revise final version.
                        \end{itemize}\\
        \hline
        October & 43 & \begin{itemize}
                            \item Revise final version.
                        \end{itemize}\\
        \hline
        November & 44 & \begin{itemize}
                            \item 01/11 Final thesis submission deadline.
                        \end{itemize}\\ 
        \hline
    % \end{tabular}}
\end{longtable}